---
title: "covid 19"
author: "Eduard Noriega"
date: "10/12/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$: Dying is independent of sex (male or female).

$H_A$: Dying is depends on sex (male or female)
  
  
Based on gender, CAT_CAT helper file will be used.

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

Based on this table above, the males tend to die more from covid-19. Where out 1411 44 died for males. 1831 females, 25 expired.

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  
Among those who died of covid-19, more males were seen to die compared to female, as shown on the data plot chart above.


  
## Numerical results  

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

64% of males are shown to have died of covid-19, which is almost twice the amount compared to females.

## Inferential results   

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value result is 0.0012 shown in chi square test, which is less than 0.05. The odds are that males in Korea are 2.3 times more likely to die from covid-19 than females.


## Conclusions

The null hypothesis is rejected, showing there is a difference between gender and dying of covid-19 in Korea. The fisher test shows a p-value of 0.0009, which is less than 0.05. 
